---
title: "Simple Linear Regression (SLR)"
author: "Davis Ogden"
date: "2024-06-16"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE,
                      include = TRUE, tidy = 'formatR',
                      fig.show = 'hold', fig.align = 'center')
```


## Simple Linear Regression (Homework 1)

### Part 1: Overview of Dataset

This homework assignment is based on a project on life expectancy research.

#### Background

Life expectancy is a measure of the average lifespan of a person born into a nation. We plan to examine the factors that can increase or decrease life expectancy and their interactions. There are almost 200 nations in the world, each with unique living conditions, economic status, and healthcare treatments. Our chosen data set contains data for 178 countries, over the 15-year period from 2000- 2015, as well as a variety of health-related and economic data about the nation for each year. It is important to note, some nations were excluded from the dataset because of repeated missing values due to difficulty finding data from smaller nations such as Togo, Vanuatu, and Cape Verde. The dataset contains four main groupings of predictor variables immunizations, mortalities, economic, and social. Data has been collected from multiple resources: the national health and economic dataset have been collected from Kaggle, and the number of medical professionals (doctors, nurses, and pharmacists, per 10,000 people in the nation) have been collected from the WHO.

``` {r Read Data}
life_data = read.csv(file = 'life_expectancy.csv')
head(life_data)
```